Characterizing dynamical transitions by statistical complexity measures based on ordinal pattern transition networks

### 2021-02-12 manuscript_Huang_rev.tex
Updated.

In the context of finite sample size effects, we note that for the particular example of the logistic map, one could easily generate realizations of much longer lengths. Our numerical results show, however, that longer realizations of a length of up to $L\leq5\cdot 10^6$ do not change the obtained entropy and SCM \emph{estimates} significantly (not shown). From a practical perspective, two aspects call for attention in interpreting the obtained entropy and SCM values quantitatively. On the one hand, due to the finite numerical precision of the computed trajectories, any chaotic regime may in fact turn out to appear as periodic (with arbitrarily long periods); this can lead to numerical artifacts that will be further discussed in Sec.~\ref{sec:transi}. On the other hand, 
%there may exist asymptotic analytical results for the case of $L\to\infty$ and $D\to\infty$, e.g., for the case of $r=4$ where one would expect $\mathcal{H}_O\to 1$ and $\mathcal{C}_O\to 0$, which substantially differ from our numerical results. In this regard, 
all estimated values of the different entropy and complexity measures always need to be considered as finite sample estimates that may well deviate from their expected asymptotic values and thereby depend on the specific values of $L$ and $D$ available for the analysis. This is, however, a feature that is also common to other existing symbolic dynamics based entropies and SCMs.

@@

\textcolor{red}{More importantly, for sufficiently large $D$ the resulting complexity--entropy pairs} essentially cover exclusively the left uprising branch of the curve for the SCM \textcolor{red}{accounting for} transition frequencies (Figs.~\ref{fig:CElogistic}(b, e) and \ref{fig:CElogistic}(c, f)), which distinctively differs from the original permutation entropy based SCM. \textcolor{red}{In this regard, one may note that this branch implies an almost linear dependency between $\mathcal{C}_W$ and $\mathcal{H}_{W}$ (Fig.~\ref{fig:CElogistic}(e)), respectively, $\mathcal{C}_{E}$ and $\mathcal{H}_{E}$ (Fig.~\ref{fig:CElogistic}(f)), which especially applies for small entropy values within an increasingly wide range as $D$ increases. This linear increase of the maximum complexity versus entropy curve for sufficiently large embedding dimension is expected from theoretical considerations~\cite{SMCbook2010}.}

>>>>

In the context of finite sample size effects, we note that for the particular example of the logistic map, one could easily generate realizations of much longer lengths. Our numerical results show, however, that longer realizations of a length of up to $L\leq5\cdot 10^6$ do not change the obtained entropy and SCM \emph{estimates} significantly (not shown). From a practical perspective, two aspects call for attention in interpreting the obtained entropy and SCM values quantitatively. On the one hand, due to the finite numerical precision of the computed trajectories, any chaotic regime may in fact turn out to appear as periodic (with arbitrarily long periods); this can lead to numerical artifacts that will be further discussed in Sec.~\ref{sec:transi}. On the other hand, all estimated values of the different entropy and complexity measures always need to be considered as finite sample estimates that may well deviate from their expected asymptotic values and thereby depend on the specific values of $L$ and $D$ available for the analysis. This is, however, a feature that is also common to other existing symbolic dynamics based entropies and SCMs.

@@

\textcolor{red}{More importantly, for sufficiently large $D$ the resulting complexity--entropy pairs} essentially cover exclusively the left uprising branch of the curve for the SCM \textcolor{red}{accounting for} transition frequencies (Figs.~\ref{fig:CElogistic}(b, e) and \ref{fig:CElogistic}(c, f)), which distinctively differs from the original permutation entropy based SCM. \textcolor{red}{In this regard, one may note that this branch implies an almost linear dependency between $\mathcal{C}_W$ and $\mathcal{H}_{W}$ (Fig.~\ref{fig:CElogistic}(e)), respectively, $\mathcal{C}_{E}$ and $\mathcal{H}_{E}$ (Fig.~\ref{fig:CElogistic}(f)), which especially applies for small entropy values within an increasingly wide range as $D$ increases. This linear increase of the maximum complexity versus entropy curve for sufficiently large embedding dimension is expected from theoretical considerations~\cite{SMCbook2010}. Some more discussions have been provided in the Supplementary Material (Sec. SM3, Fig. S6-S8). }
